---
title: "BSMM-lab-2"
subtitle: "BSMM 8740 Fall 2023"
author: "Tavish Bhandari"
date: "September 25,2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r load-pkg-data}
#| message: false
the_tate <- readr::read_delim("data/the-tate-collection.csv", ";", escape_double = FALSE, trim_ws = TRUE)
the_tate_artists <- readr::read_csv("data/the-tate-artists.csv")
```

## Exercises

### Exercise 1

The `the_tate` dataset has \_\_\_ unique artists who worked from \_\_\_ to \_\_\_. The works were acquired between the years \_\_\_ and \_\_\_.

```{r}
library(magrittr)     # the pipe
library(tidyverse)    # for data wrangling + visualization
#install.packages("tidymodels")
library(tidymodels)   # for modeling
#install.packages("gt")
library(gt)           # for making display tables
#install.packages("gtExtras")
library(gtExtras)     # helper functions for beautiful tables
install.packages("DataExplorer")
library(DataExplorer) #
```

```{r}
library(dplyr)
?summarize()
```

```{r}
library(DataExplorer)
introduce(the_tate)
```

```{r}
plot_missing(the_tate)
```

```{r}
unique_artists <- unique(the_tate[, c("artist")])
max_work_year <- max(the_tate$year, na.rm = TRUE)
min_work_year <- min(the_tate$year, na.rm = TRUE)
max_acquisition_year <- max(the_tate$acquisitionYear, na.rm = TRUE)
min_acquisition_year <- min(the_tate$acquisitionYear, na.rm = TRUE)
cat("The the_tate dataset has ", length(unique_artists$artist), " unique artists who worked from", min_work_year," to ",max_work_year, ". The works were acquired between the years", min_acquisition_year, " and ",max_acquisition_year,"\n")

# Can use ?stringr::str_glue
```

### Exercise 2

```{r}
library(tibble)
library(dplyr)
DataExplorer::introduce(the_tate)
DataExplorer::introduce(the_tate_artists)
DataExplorer::plot_missing(the_tate)
DataExplorer::plot_missing(the_tate_artists)
works_with_missing_year <- sum(is.na(the_tate$year))
the_tate|>dplyr::filter(is.na(year))|>dplyr::distinct(artist)
artists_with_missing_date <- the_tate |>
  filter(is.na(year)) |>
  distinct(artist) |>
  nrow()
artist_missing_year_count <- the_tate|>dplyr::filter(is.na(year)) |> 
  group_by(artist) |>
  summarise(missing_year = sum(is.na(year))) |> 
  arrange(desc(missing_year)) |> 
  as_tibble()

print(artist_missing_year_count)

artists_with_missing_year <- nrow(artist_missing_year_count)

print(artists_with_missing_year)

artist_missing_year_count <- artist_missing_year_count |> # Calculate the percentage of missing 'year' values for each artist #
  mutate(percentage = (missing_year / works_with_missing_year) * 100)

artist_missing_year_count <- artist_missing_year_count |>
  mutate(cumulative_percentage = cumsum(percentage)) # Calculate the cumulative percentage of missing 'year' values # 

artists_to_resolve_50_percent <- min(which(artist_missing_year_count$cumulative_percentage >= 50))

print(artists_to_resolve_50_percent)
```

How number of works with missing date is 5397.

The number of artists whose works have missing year is 461.

It would require resolving missing year data for only 11 artists resolve resolve at least 50% of the missing data.

The missing year data likely to be classified as MAR.

### Exercise 3

```{r}
# Calculate the count of works for each artist in the 'the_tate' dataset #
artist_work_counts <- the_tate |>
  group_by(artist) |>
  summarize(title = n()) |>
  arrange(desc(title))

# Top 10 artists with the most works #
top_10_artists <- artist_work_counts |>
  slice_head(n = 10)

print(artist_work_counts)
print(top_10_artists)
```

The artist with the most works in the Tate collection is Turner,Joseph Mallord William.

The artist with the tenth-most works in the Tate collection is Warhol,Andy.

### Exercise 4

```{r}
library(dplyr)
library(gt)
library(gtExtras)
total_works <- nrow(the_tate)
artist_work_counts %<>%mutate(percentage=(title/total_works)*100)
print(artist_work_counts)

# gt table creation #
table <- artist_work_counts %>%
  gt() %>%
  fmt_number(
    columns = c(title, percentage), # Format both title and percentage columns
    decimals = 2
  ) %>%
  
#Display the formatted table using print #
print(table)
```

The artist with the greatest number of works in the Tate collection represent 56.9 % of the total number of works

### Exercise 5

```{r}
total_rows <- total_works

# Select only the columns for artist and title to count distinct combinations #
distinct_artist_title_pair <- the_tate |> select(artist,title) |> distinct()

distinct_count <- nrow(distinct_artist_title_pair)
# Count and print the number of duplicated artist-title pairs
duplicated_count <- total_rows - distinct_count
print(duplicated_count)
```

There are 23705 duplicate artist-title pairs

### Exercise 6

```{r}
library(tidyr) # Load tidyverse library 
the_tate <- the_tate |>
  mutate(area_cmsquare= width * height/10000)
selected_artworks <- the_tate |>
  select(artist, title, area_cmsquare) |>
  drop_na() # Remove rows with containing NA values#
ordered_artworks <- selected_artworks |>
  arrange(area_cmsquare)
largest_artwork <- ordered_artworks |>
  slice_tail(n = 1)
smallest_artwork <- ordered_artworks |>
  slice_head(n = 1)
print(smallest_artwork)
```

The artist with the largest work in the tate collection is Therrien, Robert.

The artist with the smallest work in the collection is Mesens, E.L.T.. The smallest work has area 0.0237 $\text{cm}^2$

### Exercise 7

```{r}
library(dplyr)
gender_grouped<- the_tate |>
  dplyr::left_join(the_tate_artists, by = c("artist" = "name")) |>   tidyr::drop_na(gender) |> group_by(gender) 
print(gender_grouped) # display the gender grouped table
```

...

### Exercise 8

```{r}
library(dplyr)
library(readr)
data <- read_csv("data/SPX_HistoricalData_1692322132002.csv") 
data <- data |> # Add a column for the year of transaction #
  mutate(Year = lubridate::year(as.Date(Date, format = "%m/%d/%Y")))
data<- data |>
rename("close"=`Close/Last`)
data <- data |>
  mutate(rd = log(lead(close) / close))
data<- data |>
  mutate(vard = rd^2)
summary_data <- data |> #summarise the data to group by year #
  group_by(Year) |>
  summarize(
    Annual_Return = (exp(sum(rd, na.rm = TRUE)) - 1)*100, 
    Annual_StdDev = sqrt(sum(vard, na.rm = TRUE))*100,
    .groups = "drop"  # Drop grouping after summary
  )
print(summary_data)
```

The annual return in the SPX price in 2020 was -13.98%.

The corresponding price volatility was 34.70%.

### Exercise 9

```{r}
library(dplyr)
library(readr)
period_return <- prod(1 + summary_data$Annual_Return) - 1
period_volatility <- sqrt(sum(summary_data$Annual_StdDev^2))
summary_rows <- tibble::tibble(
  Year = as.character("Period"),  # Ensure "Year" is character type
  Annual_Return = period_return,
  Annual_StdDev = period_volatility
)
summary_data <- summary_data |>
  mutate(Year = as.character(Year))

# Combine data
summary_data <- bind_rows(summary_data, summary_rows)
print(summary_data)
```

The period volatility was 48.77%

### 
